### 2024-11-17 paper.tex
added outline to paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

	\title{Integration of 75MW Solar PV Plant: Transmission System Design Analysis}

	\author{\IEEEauthorblockN{Brent Dickinson}
		\IEEEauthorblockN{Tianci}
	}

	\maketitle

	\begin{abstract}
		This report presents the design and analysis of transmission system modifications required to integrate a new 75MW solar PV plant while addressing existing system reliability concerns. The study evaluates various transmission line and transformer options to determine the most cost-effective solution that maintains system stability under both normal and N-1 contingency conditions.
	\end{abstract}

	\section{Introduction}
	\begin{itemize}
		\item Project overview and objectives
		\item System requirements and constraints
		\item Design methodology
	\end{itemize}

	\section{Initial System Analysis}
	\subsection{Base Case Evaluation}
	\begin{itemize}
		\item Existing system configuration
		\item Current system performance metrics
		\item Identification of existing violations
	\end{itemize}

	\subsection{Contingency Analysis Results}
	\begin{itemize}
		\item N-1 contingency findings
		\item Critical violations and their implications
		\item Key problem areas identified
	\end{itemize}

	\section{Design Solutions}
	\subsection{Design Options Considered}
	\begin{itemize}
		\item Transmission line alternatives
		\item Voltage level considerations
		\item Transformer placement options
	\end{itemize}

	\subsection{Technical Analysis}
	\begin{itemize}
		\item Power flow studies
		\item Contingency analysis results
		\item System performance improvements
	\end{itemize}

	\section{Economic Analysis}
	\subsection{Cost Components}
	\begin{itemize}
		\item Fixed costs (lines, transformers, substations)
		\item Variable costs (distance-based)
		\item Loss reduction benefits
	\end{itemize}

	\subsection{Solution Comparison}
	\begin{table}[htbp]
		\caption{Cost Comparison of Design Options}
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				\textbf{Design Option} & \textbf{Capital Cost} & \textbf{Loss Savings} & \textbf{Net Cost} \\
				\hline
				Option 1 & & & \\
				\hline
				Option 2 & & & \\
				\hline
			\end{tabular}
			\label{tab:costs}
		\end{center}
	\end{table}

	\section{Recommended Solution}
	\begin{itemize}
		\item Final design selection
		\item Technical justification
		\item Economic justification
		\item Implementation considerations
	\end{itemize}

	\section{Conclusion}
	Summary of key findings and recommendations

	\begin{thebibliography}{00}
		\bibitem{b1} PowerWorld Corporation, "PowerWorld Simulator Manual," 2024.
		\bibitem{b2} IEEE Standard 1547-2018, "IEEE Standard for Interconnection and Interoperability of Distributed Energy Resources with Associated Electric Power Systems Interfaces."
	\end{thebibliography}

\end{document}